---
title: "priors"
author: "Mark Hagemann"
date: "November 2, 2016"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, results = "asis",
               warning = FALSE, message = FALSE)
opts_knit$set(root.dir = "..")
```

```{r}
library(ProjectTemplate)
load.project(override.config = list(cache_loading = TRUE, munging = FALSE))
library(visreg)
library(stargazer)

doctype = "html"
```


The hydroSWOT dataset allows us to mine for the following prior distributions:

### Ao (conditional on W statistics) -- 

- lognormal with 
- mu = 0.8549474 + 1.3925854 * lwbar - 1.7249043 * lwsd, 
- sigma = 0.8229

```{r}
fromcache("alm")
coefs <- coef(alm)
nc <- length(coefs)

ggTermPlot(alm, xvar = names(coefs)[2], data = alm$model)
ggTermPlot(alm, xvar = names(coefs)[3], data = alm$model)

stargazer(alm, type = doctype)
```

The standard deviation comes from the GCV: 

```{r, echo = TRUE}
sqrt(gcv(alm))
```


### b (conditional on sd(logW))

- truncated normal with mean 0.0231 + 0.454 * sd(logW), variance 0.00975

```{r}
fromcache("blm")
fromcache("bpreddat")
coefs <- coef(blm)
nc <- length(coefs)

vars <- all.vars(blm$call$formula)[-1]

ggTermPlot(blm, xvar = vars[1], data = bpreddat)

stargazer(blm, type = doctype)
```

The standard deviation comes from the GCV: 

```{r, echo = TRUE}
sqrt(gcv(blm))
```



### AHG error: (don't actually need this--see math)

- relate to cor(logW, logQ)
- $var(e_{ahg}) = var(logW) - b^2var(logQ)$

### logQ

- hyperprior on mean
    - centered at logQWBM
    - spread?
- hyperprior on variance
    - get from database
    - Here regulation structures, land-cover, slope?, climate would be useful
    - relate to var(logW)
    - sd(logQ) is normal with mean 0.72469 + 0.5804 sd(logW) + 0.1509 sd(H), stdev 0.3587.
    - all units in SI, all lograithms lm. 

```{r}
fromcache("qsdlm")
coefs <- coef(qsdlm)
nc <- length(coefs)

ggTermPlot(qsdlm, xvar = names(coefs)[2], data = qsdlm$model)
ggTermPlot(qsdlm, xvar = names(coefs)[3], data = qsdlm$model)

stargazer(qsdlm, type = doctype)
```


The standard deviation comes from the GCV: 

```{r, echo = TRUE}
sqrt(gcv(qsdlm))
```



### logn

Prior is normal(-3.5, 1). Resulting in a lognormal(-3.5, 1) prior on n. That looks like:

```{r}
ns <- seq(0.001, 0.3, length.out = 10000)
plot(ns,  dlnorm(ns, -3.5, 1), type = "h")
```


### Manning error

I haven't found a good source for Manning's error statistics. But Dingman (1997) gives RMSE of log10(Q) for a couple modifications of Manning's equation. This is given as about 0.25. 



### logQc

We still don't really know what Qc represents, physically. So hard to constrain at this point. But best guess is it's near mean(logQ)

### logWc

Also not sure on this one. We'll know more when we get an AMHG validation set.